\begin{preface}
\begin{fancyquote}{\textit{Tim Berners-Lee}}
``I found myself answering the same questions asked frequently of me by different people. It would be so much easier if everyone could just read my database.''
\end{fancyquote}

\noindent Solid is an amazing technology that has the potential to change the web as we know it. The last year, I have witnessed amazing progress being made in this domain, and I am truly honored that I had the opportunity to be a part of it. It has been quite the roller coaster, combining my thesis research with a semester spent abroad on an Erasmus exchange, but it was well worth the effort.\\

\noindent There are an endless amount of people I want to thank for supporting me through this period, especially my friends and family.\\

\noindent First and foremost, my most sincere gratitude goes to my girlfriend Elise, who had to endure endless streams of complaints when I was stuck somewhere, and screams of joy when I made some progress. Without her support, this text would not be the same.\\

\noindent Secondly, I owe an immense amount of gratitude to my supervisors and assistant-supervisors. 
In particular, I want to thank Bert Lagaisse for allowing me to combine my thesis while studying in Hungary, which was not a trivial task. Your feedback and guidance taught me a lot and greatly improved the quality of my work. I also want to thank Emad for the great advice and encouragement every week. Your advice has inspired me many times and your kind words have always encouraged me. Finally, I also want to explicitly thank prof. Vincent Naessens, prof. Wouter Joosen and Dimitri Van Landuyt for their valuable feedback and suggestions.\\

\noindent Thirdly, I want to thank my family for supporting me not only through this thesis, but throughout my whole study career. Without their support, I would not have become the person I am today.\\

\noindent Lastly, I want to thank ... and my friends ..., ... and ... for not only supporting me throughout the whole year, but also for proofreading and improving this text. 

\end{preface}