{\small Solid is a W3C specification that aims to decentralize the web using personal online datastores (pods). Pods are decentralized repositories where data is held for different applications. A challenge here is secure and protected data aggregation across pods. Such aggregations present privacy risks as well as scalability issues. This dissertation introduces a middleware that attempts to partially (at the server level) solve these problems, thanks to the introduction of privacy filters and a new access token mechanism that supports decentralized access token delegation.
Privacy filters allow to achieve more granularity in the privacy of shared resources. This is realized by having the middleware automatically select transformations to be performed when a resource is requested, based on a number of contextual parameters. Subsequently, to obtain a token mechanism that supports decentralized access token delegation, this dissertation examines the use of macaroons within Solid. Macaroons not only support decentralized access token delegation, but are also more efficient to generate and verify. In addition, they allow third-party attestations, which is a useful property for enabling group vaults, or pods shared by a number of people.
The middleware presented in this dissertation was evaluated based on three use cases. For privacy filters, experiments have demonstrated that the overhead of the middleware is tolerable for smaller resources with an overhead of around 50\%. However, larger resources (around 3MB) have an overhead that becomes five times the original duration of the request. Performance tests focusing on the throughput of token generation and verification have shown that macaroons have a throughput that is seven and eleven times greater than the DPoP (ES256) system, respectively. Finally, there are also theoretical gains for decentralized token delegation.}
